### 2022-06-08 PaperAnalysis.Rmd
Heterogeinity analysis explained

---
title: "Causal Analysis Paper"
author: "Helena Bakic, Anna-Lisa Wirth, Ángela Duarte Pardo"
date: "5/23/2022"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dagitty)
library(ggdag)
```

\newcommand{\indep}{\perp \!\!\! \perp}

# Background

Abortion law and the negative consequences of the lack of safe abortions are highly contentious policy issues that have been making headlines this year. Further, the Sustainable Development Goal indicator 3.1 of the United Nations to reduce Maternal Death to less than 70 per 100000 life birth by 2030 (https://www.who.int/data/gho/indicator-metadata-registry/imr-details/26) is relying on access to safe and legal abortions. In the paper we are analysing the authors are testing whether and if so to what extent abortion law as a proxy for access to safe abortions causally effects maternal mortality rates in over 160 countries around the globe. Establishing a causal relation could strengthen the implementation of abortion liberalisation policy. This is especially needed in a time where several states already withdrew or are withdrawing access to safe abortions on many grounds such as Poland (https://www.amnesty.org/en/latest/news/2021/10/poland-a-year-on-abortion-ruling-harms-women/) and the USA (https://www.euronews.com/2022/05/03/us-supreme-court-moving-to-overturn-key-abortion-rights-law-says-leaked-report). While on the other hand there are developments like for instance in France to expand abortion rights (https://www.hrw.org/news/2022/03/01/france-expands-abortion-access-two-key-moves) in 2022. So does abortion law flexibility cause maternal mortality? We will be taking a close critical look at a paper that tries to establish exactly that and also run our own data analysis to address this question.

# Description of the original paper

In "Abortion laws reform may reduce maternal mortality: an ecological study in 162 countries," @Latt2019 present the results of a study that aims to assess whether maternal mortality is affected by the flexibility of abortion laws. Abortion law flexibility was measured as a sum of the number of reasons for which abortion is legally allowed (ranging from 0 to 7) and was included in models as both an ordinal variable and a binary variable (inflexible vs. flexible abortion laws). Abortion laws were considered inflexible if there were less than 3 legal grounds for abortion. Maternal mortality ratio (MMR) was measured as the number of maternal deaths per 100,000 live births. Authors used fixed effects linear regression models to estimate the relationship between abortion law flexibility and MMR, while controlling for GDP per capita and time in five-year intervals. Additional analysis included also female primary education completion as a control variable as it had over 50% of missing data; however, results did not change. Results indicated that maternal deaths reduce by 45 per 100,000 live births (95% CI: -64, -26) with flexible abortion laws. The authors conclude that abortion law reform in countries with inflexible abortion laws may reduce maternal mortality.

#  Methodology

## Theorethical estimands

Theoretical estimand is a key part of a research question, as it describes the quantity of interest independently of statistical models and clearly describes the research goal [@Lundberg2021]. In the current paper, the theorethical estimand was not specifically stated; however, we tried to deduce the desired estimand from the research goal and the estimation strategy authors used.

The authors state that they aim to “estimate the association between the flexibility of abortion laws and maternal mortality at a country level [between 1985 and 2013]” [@Latt2019, pg. 2]. Furthermore, they estimated the model using linear regression, and the quantity they have estimated was the regression coefficient of MMR on flexibility of the law. Therefore, we we believe that the authors wanted to assess the average treatment effect (ATE), that is, the difference in the number of maternal deaths that would occur within a country if abortion law was flexible ($MMR_i$ = 1) versus inflexible ($MMR_i$ = 0):

$$
ATE = \frac{1}{n}\sum_{i=1}^{n}(MMR_i^1 - MMR_i^0) = \frac{1}{n}\sum_{i=1}^{n}\Delta MMR_i = \mathbb{E} [\Delta MMR]
$$

This estimand looks at counterfactual outcomes for all countries, regardless of the fact that the particular country would in reality receive treatment or not. Country $i$'s MMR when abortion laes are flexible is denoted as $MMR_i^{(L=1)}$ or shortly $MMR_i^{(1)}$, and as $MMR_i^{(L=0)}$ or $MMR_i^{(0)}$ when abortion laws are inflexible. Individual causal effect for the country $i$ is then the difference between mortality when abortion laws are flexible vs. when they are inflexible, or $ΔMMR_i = MMR_i^{(1)} - MMR_i^{(0)}$. Individual effects averaged across all of the countries in the world would then give the wanted causal quantity. This is equal to the expectation of the effect of flexible abortion laws on MMR of a randomly selected country.

As ATE reflects the effect of the treatment for the entire study population, it can be particularly useful for scoping the benefit of population-wide policies. However, as there will be inevitable self-selection to treatment in case of abortion policies -- some countries will choose to have more flexible laws and some will choose less flexible laws -- there are other estimands that could potentially be interesting for researchers in this particular case.

In fact, one interpretation of the results that authors give is that "it is possible that maternal mortality would reduce in [countries with low abortion flexibility scores] if legal, safe abortion was more readily available" [@Lundberg2021, pg. 6, para. 2]. This would indicate that they might actually be interested in the ATU - the average effect of the treatment on the untreated, that is, for the population of units that did not receive the treatment:

$$
ATU = \mathbb{E}[MMR^{(1)} - MMR^{(0)} \mid L = 0]
$$

ATU shows how the outcome would differ if untreated units were actually treated, in other words, the effect of expanding the treatment. This could potentially be the most useful estimand in this study, as it could be used in considering the expansion of a medical practice to units that were previously untreated [@greifer2021].

Another potentially interesting estimand is the ATT, that is, the effect of the treatment for the population of units that actually received the treatment:

$$
ATT = \mathbb{E}[MMR^{(1)} - MMR^{(0)} \mid L = 1]
$$

ATT shows how the outcome would differ if treated units were actually not treated, in other words, the effect of the withholding the treatment from the ones who would otherwise receive it [@greifer2021]. In light of recent policy changes and reductions of abortion flexibility laws in the USA, ATT could be particularly relevant.

In an ideal (in terms of causal identification) scenario -- randomised control trials (RCTs) -- all three estimands (ATE, ATU, ATT) will be the same because we assume that:

* The baseline of the treated group is equal to the baseline of the untreated group, that is, countries that have flexible abortion laws would do *as bad* (in terms of MMR) as countries that have inflexible abortion laws, if they didn't have flexible abortion laws:

$$
\mathbb{E}[MMR^{0} \mid L = 1] = \mathbb{E}[MMR^{0} \mid L = 0]
$$

* The treatment effect will not be different in the treated and the untreated group, that is, countries that have inflexible abortion laws would do *as good* as countries with flexible abortion laws, if they made the laws more flexible:

$$
\mathbb{E}[MMR^{1} \mid L = 1] = \mathbb{E}[MMR^{1} \mid L = 0]
$$
These assumptions are commonly violated in observational studies, as we will show in the next section, therefore it is to be expected that the three estimands (ATE, ATU, ATT) would be different in the current study.

## Identification

For each country $i$ it is not possible to observe both potential outcomes, so a theoretical estimand is not an empirical quantity. Emprirical estimand is a translation of the theoretical, unobserved quantity, into observable data. This link is often formalised using Directed Acyclic Graphs (DAGs) that formally state model assumptions under which a theorethical estimand can be calculated [@Lundberg2021].

The current paper proposes that flexibility of abortion laws causes maternal mortality, and that GDP and female primary education completion are confounders of this relationship -- that is, they cause both whether or not a country will have flexible abortion laws and how big MMR will be. The authors also propose that the causal mechanism by which this relationship comes about is that the women in countries where restrictive laws are in place are less able (or unable) to access reproductive health care and safe abortion services, which increases the proportion of unsafe abortions, and in turn, maternal deaths. The model can be seen in the figure below:

<center>

![Figure 1.: Causal model as recreated from @Latt2019 ](Figures/dag1.png)

</center>

This model has several practical implications:

* The effect that we are trying to estimate, the effect of abortion law on MMR, has 2 "paths": Law &rarr; Access to safe abortion &rarr; MMR and Law &rarr; Access to safe abortion &rarr; Number of safe abortion &rarr; MMR;
* Even though Access to safe abortion and Number of safe abortions are unmeasured, the relationship between Law and MMR can be estimated (but we won't know how much of that relationship is due to the Number of unsafe abortions); 
* Female primary education completion and GDP confound the Law &rarr; MMR relationship;
* Because Female primary education and GDP are measured, then $MMR^{(1)}, MMR^{(0)} \indep Law \mid Female \ primary \ education, GDP$.

However, there are several potential identification issues that could influence the possibility to estimate the effect of abortion law flexibility on MMR, namely; unmeasured confounding and (non-random) data missingness. These are discussed in detail below.

### Unmeasured confounding

What makes the estimation of ATE plausible is “uncoupling” or *d-separation* of assignment of units of analysis (in this case, countries) into treatment and control conditions. In randomized trials, this is achieved by randomly assigning treatment to unit analysis (Figure 2a). Randomization:

* Identifies the effect of Abortion law on MMR;
* Generated approximate balance in counfounders, even if a counfounder is unobserved;

which means that the causal effect of Law on MMR can be estimated. Conditioning on confounders is still (statictically) desirable, as it will increase the precision of the effect estimate.

```{r, echo=FALSE,out.width="49%",out.height="49%",fig.show='hold', fig.align='center', fig.cap="Figure 2. a) Randomisation removes inbound connections from confounders to treatment, b) Without controling for all counfounders, the causal effect can't be estimated"}

knitr::include_graphics(c("Figures/dag2a.png","Figures/dag2b.png"))
```

However, in observational studies, the causal effect is identified only when we control for all confounders. If there is unobserved confounding (Figure2b), the causal effect can't be identified as:

$$
MMR^{(1)}, MMR^{(0)} \indep Law \mid Female \ primary \ education, GDP, Unmeasured \ confounders
$$

Given that unmeasured confounding is a crucial issue in effect identification, we have considered possible omitted variables and their role in the original model:

* Health expenditure: health expenditure likely determines the access to safe abortions and women’s trust in medical institutions. The original paper found that GDP and health expenditure are so closely related that they used GDP as a proxy. We will keep this assumption.

* Religious views in society and level of secularism: The power of religious institutions and the views they hold in regard to the acceptability of abortion is likely to influence the level of maternal mortality and the flexibility of abortion law. If for instance a Christian Catholic institution has a lot of power in society and is very traditional this likely will influence the flexibility of the law but also the individual’s choice whether to get a clandestine abortion outside of the health system or whether she will even have the possibility to acquire a safe abortion in a health institution. However, this confounder would be extremely difficult to measure as there is no systematic recording of religious power and the views different institutions hold for all the studied countries.

* Party in power politically: The political convictions and policies of political parties in power most likely will influence the abortion law. However, it is unlikely that it will directly cause mortality. Therefore, it is not a confounder in this model.

* State of human rights and gender equality: State of human rights and gender equality in a country could influence the abortion law flexibility and female primary education completion. However, we think it is unlikely that they will impact MMR; therefore, it would not be necessary to control for them, as long as female primary education is controlled for.

Lack of treating the missingness with female literacy

Comparing the rate of MMR and flexibility score for countries with recorded female primary completion vs countries without recorded female primary completion levels we could check whether there is a selection bias and then generalization of effects of this model would not be possible. If we suppose that female literacy is a collider and a confounder, we cannot condition on it without creating selection bias {@elwert2014endogenous, p.38}. As this done in the baseline paper, we would have to select strategies to correct this. Otherwise, it would stay d-connected, so statistically dependent. 
We see issues with potential interconnectedness and association of female literacy with abortion laws and access to abortion. If there is no school that can teach women how to read most likely there is no clinic that can perform a safe abortion, further women are unlikely to be informed about such services. While the issue of women’s information is taken up in the original paper and the lack of knowledge about abortion services even stays a problem for countries with high flexibility scores, low female primary completion here used as a better recorded proxy for female literacy is likely to cause access to abortion. So, it is a post-treatment confounder that potentially influences the outcome of MMR and abortion laws or the original variable of interest - access to safe abortions - it is a proxy for. The exact path between those variables should be analysed. The original paper suggested 2 models for this:

Model 1 included all the countries and no female literacy proxy:

```{r}
dag  <- dagify(MMrates ~ laws,
               MMrates ~ GDP,
               laws ~ GDP
               )
ggdag(dag) +  theme_dag_blank()
```

```{r}
dag  <- dagify(MMrates ~ laws,
               MMrates ~ GDP,
               MMrates ~ literacy,
               laws ~ GDP,
               laws~literacy
               )
ggdag(dag) +  theme_dag_blank()
```

However, it is also possible that female literacy has no effect on abortion law and actually only has an effect on the outcome, MMR. MMR is a common effect of female literacy and flexibility score (abortion laws). Then instead of being a causal effect to abortion law female primary completion would be an associative risk factor.{@10.2307/20485961, p.3}. In this case the only connection would be if we conditioned on MMR we would create a spurious relation however the effect of literacy and law on each other would be zero. Yet, we do not condition on the outcome hence no spurious relationship should be created.

```{r}
dag  <- dagify(MMrates ~ laws,
               MMrates ~ GDP,
               MMrates ~ literacy,
               laws ~ GDP
               )
ggdag(dag) +  theme_dag_blank()
```

However, it is difficult to disentangle the relationship as lower female literacy will also lower access to safe abortions regardless of the flexibility of the law as women are likely less informed about their health care options which in turn increases female mortality.

Testing for those different possible scenarios would be possible with sensitivity analysis or inverse probability weighting. We would need to control for the differences in the sets of considered countries (female lit = 1, recorded) and not considered countries (female literacy = 0, not-recorded) and see what other differences there are or trends in the groups other than female literacy being observed. While the authors drew two models they did not make this clear comparison in the paper. This additional analysis would help us to gain insight into whether recorded primary completion influences female mortality and to what extent. Further, it should be checked whether primary completion is a function of the state of human rights and GDP. The general state of human rights and equality could also influence abortion law and thus should be studied in order to make policy recommendations. However, a coded human rights and equality score that would be usable for causal analysis would need to be found. The IDEA democracy indices could be considered as a proxy for this (https://www.idea.int/data-tools/tools/global-state-democracy-indices).

A possible model for future research could be:

![Figure 1.: Suggested causal model for future research ](Figures/DAG_4.png)

```{r}
 testImplications <- function( covariance.matrix, sample.size ){
	library(ggm)
	tst <- function(i){ pcor.test( pcor(i,covariance.matrix), length(i)-2, sample.size )$pvalue }
tos <- function(i){ paste(i,collapse=" ") }
implications <- list(c("A","GDP","E","HRE"),
		c("A","GDP","E","L"),
		c("A","HRE","E","L"),
		c("E","L","HRE"),
		c("GDP","L","HRE"),
		c("HRE","MMR","A","E","GDP"),
		c("HRE","MMR","E","GDP","L"),
		c("L","MMR","A","E","GDP"),
		c("L","MMR","A","E","HRE"))
	data.frame( implication=unlist(lapply(implications,tos)),
		pvalue=unlist( lapply( implications, tst ) ) )

}
```

HRE: Human Rights and Equality
GDP : GDP
L : Abortion Law
A : Access to Abortion
E: Female Education
MMR : MMR

Sensitivity Analysis

In the sensitivity analysis we would check which part of the total effect can be explained by the abortion laws and which can be explained by female literacy. Additionally, we check how much the effects would need to change to cancel one another.For instance, the effect for flexibility score alone we can check when we control (block) the primary completion rate.
Sensitivity analysis has 2 essential parts to it:
1)describing the type of unobserved confounders
Unmeasured variables: religious institutions, contraceptive prevalence, fertility rate 
2) plausibility that problematic confounding might exist
	Based on 
	*research design, 
  *expert knowledge and 
  *data-generating process

Generally, sensitivity analysis is still underutilized in social science and policy research papers {@cinelli2020making, p. 40}. However, it is a very useful tool to test robustness of findings, i.e. how sensitive my effects and effect sizes are to other variables that influence the treatment or outcome studied. Therefore, we are proposing to add a sensitivity analysis to the research design and also apply it in the second part of our report. This method gives us two insights: 1) the robustness value: how robust the coefficient is to confounders influencing treatment and outcome and 2) the proportion of the outcome uniquely explained by the treatment and how strong a confounder would have to be to eliminate the effect of the treatment{@cinelli2020making, pp.40/41}.

### Missing data

Hernan et al. are outlining in their paper why data is missing, to restrict the analysis on countries where data are complete would bias it. Countries with missing data on female literacy are still included {@10.2307/20485961, p. 5}. Female primary completion is only available for 50 percent of countries and female literacy only for 10 percent. This missingness of data introduces selection bias to this specific analysis the auhtors did in the original paper. As the likely selection bias in this study cannot be avoided due to the limitations on data availability it needs to be corrected if possible, in the analysis {@10.2307/20485961, p. 7} this could be achieved through generalization of inverse probability weighting. For this however we would need to categorize developing countries in one group where some have recorded female literacy which then gets multiplied by the number of countries not considered due to missing data. Inverse probability weighting would work as the score for female literacy is not zero for the countries where it is not recorded.  However, a lot of assumptions would need to be made to apply this method. The assumption that the filler-country would be close enough to the unobserved countries is by nature untestable and therefor it is a strong assumption.

## Empirical estimation

To estimate the effects of Abortion law flexibility on MMR, the authors of the paper fitted a fixed effects multiple regression model, in which they control for GDP, time, and countries, while trying to estimate the regression slope of flexible abortion law as a causal effect that is being studied:

[move Angelas equation here?]

This approach has been called "the workhorse model" used by researchers to estimate causal effects [@Aronov2016]. What fixed effects models do is that they remove all the unmeasured but group-specific confounding effects related to the unit that is being controlled. Controlling for unobserved counfounding related to countries and time is a common approach in studies such as these. However, this approach can have subtle consequences that limit the generalizibility of the findings.

@Aronov2016 showed that multiple regression estimates are actually weighted averages of unit-specific contributions. In cases when treatment effects are heterogeneous, in the sense that the effect of abortion law flexibility is not the same for every country, $\beta_L$ will not reflect ATE. Rather, it will be more strongly influenced by units that have more variability: many observations, large variations in abortion law flexibility and MMR, and that are hard to predict from country and time fixed effects and other controls (GDP, female primary education completion). This means that the estimated effect will be actually valid for a *sub-sample* of the population that is being studied. @Aronov2016 suggest that the *effective* sample, the sub-sample that has the strongest weights in the analysis should be reported explicitly so that it is more clear to which cases the effects generalise.

G-separation is a method that enables the unbiased estimation of marginal causal effects by using counterfactuals. It  responds to two of the effect identification issues encountered in observational studies: the missingness of counterfactuals (because it is impossible to observe participants under both treatment and non-treatment) and heterogeneity of effects (variation of effects across individuals).[@snowden2011implementation]

G-separation calculation has two steps: 
1. Fitting a Q-model, that is, a traditional regression of the outcome on the treatment and  relevant covariates.  In our case, we used the traditional regression fitted in the original paper as Q model:

$$
MMR=\beta_0 + \beta_1 L+\beta_2 GDP +\beta_3 P+ \beta_4 C
$$
In the formula, P and C respectively account for 3 year periods and the country specific effects. 
2. The Q-model is used to predict counterfactual outcomes for all the observations under both treatment settings (L=1 for flexible abortion laws and L=0 if not). More specifically, we plugged both treatment regimes in the the regression fit and obtained two predicted outcomes: $Y_{L=0}=Y_0$ for the unexposed and $Y_{L=1}=Y_1$ for the exposed. Both outcomes are then used to calculate the average $Y_1-Y_0$, which corresponds to the ATE.

With this intervention, the missingness problem is resolved because counterfactual outcomes for each observation are generated. G-separation thus provides a full dataset "free of confounding under causal assumptions."[@snowden2011implementation].

Furthermore, G-separation allows for the calculation of both ATT and ATU by restricting the ATE estimation to the treated and the untreated respectively[@Wang].

# Analysis
Our empirical exploration had two main objectives: 1. to reproduce the methodology of the original paper with updated data on the same, 2. to compare the result thus obtained with some other methods that address the assumptions not met by the original paper.

* The updated dataset
Although the datasets used in the original paper were hosted in public websites, the currently available version is not exactly the same. Since both the abortion policies and the maternal mortality sets are now hosted in interactive webpages, it was hard to access the aggregated and historical data from both abortion policies and maternal mortality rates between 1985-2013. Despite these restrictions, we collected data on all the relevant variables from 2000 to 2017. The maternal mortality data was accessed in the World Bank Data repository[@MMR], and the global abortion policies were found in the UN Department for Economic and Social Affairs archive [@WorldAbortionPolicies].

As in the original dataset, the Maternal mortality ratio (MMR) was measured as the number of maternal deaths per 100,000 live births and legal grounds for abortion where coded as binary variables ("1" for countries that legally allow for abortion on that ground; "0" for those that don't). We also calculated an aggregated and ordinal flexibility score for abortion by adding the legal grounds together.

Unlike the authors, our binary flexibility of abortion policies variable is slightly more stringent: since the median of the abortion grounds calculated every 3 years is approximately 4 in average, we determined that abortion laws are inflexible if countries allow for less than 4 legal grounds for abortion.

*[Add table showing median per period.]*

* Regression ATE
In the first part of our exploration, we followed the methodology of the original paper and used fixed effects linear regression models to estimate the relationship between abortion law flexibility and MMR. We controlled for GDP per capita, time in three-year intervals, and country-specific effects through a factor country-name variable. According to our results, flexible abortion laws reduce maternal deaths by 57 per 100.000 live births [ADD TABLE]. This result is close enough to the original result, which found that maternal deaths reduce by 45 per 100,000 live births (95% CI: -64, -26) with flexible abortion laws. 
We also conducted added "female primary education completion" as a control variable, and found very similar result: the effect of abortion law flexibility on maternal mortality appears to be a reduction in the mortality rate of 53 deaths per 100.000 live births.

* Heterogeinity of effects

Following the approach of @Aronov2016, we have reproduced the weights that multiple regression analysis uses to caluclate the effects by 1. regressing the confounders and the fixed effects for countries in the analysis and the time period on the abortion law flexibility (measured as a score from 0 to 7) 2. squaring the weights.

As expected, our analysis showed that countries have different weights in the calculated effect. The nominal sample, the full population of the countries that have data for both the treatment and the outcome, is shown on Figure Xa and it includes almost all countries in the world. However, the effective sample, defined here as the countries that contribute over half (51%) of the weight used to construct the estimate of the abortion law flexibility on MMR, consists of only 14 countries (Figure Xb).

```{r, echo=FALSE,out.width="49%",out.height="49%",fig.show='hold', fig.align='center', fig.cap="Figure X. a) Nominal sample - all the countries in the analysis b) Effective sample - countries that contribute over 51% of the effect of abortion law flexibility on MMR"}

knitr::include_graphics(c("Figures/full_sample.png","Figures/effective_sample.png"))
```

As expected, countries with more data and higher variability in abortion law flexibility and MMR contribute more to the effect, as can be seen from the Figure X below.

```{r, echo=FALSE,out.width="49%",out.height="49%",fig.show='hold', fig.align='center', fig.cap="Figure X. a) Countries that contribute the most to the effect of abortion law flexibility on MMR vs. b) Countries that contribute the least"}

knitr::include_graphics(c("Figures/most.png","Figures/least.png"))
```

* G-separation ATE, ATU, ATT

* Unmeasured confounding

# Conclusion 
The policy conclusion the original paper draws is justified. However, we could expand upon the methodology to get a more precise result and consider possible models as the limitation of models is not sufficiently justified in their paper. There are several ways to amend or extend the original research design. We presented suggestions in the theoretical part of our report like applying sensitivity analysis. We generated partial insights with our own data analysis. The causal relation between abortion law and maternal mortality ratios could be established. However, the proportion of the effect remains to be studied with an extended methodological toolbox.

# github link to our analysis project: https://github.com/aduarte56/AbortionLaws

# Contributions

We worked together as a team where we all worked on all parts and engaged with our colleagues' ideas through out. Each of us took the lead on parts of the report as detailed below:

Helena Bakic : Empirical estimation,Identification, Heterogeinity of effects, Analysis
Anna-Lisa Wirth (213286@mds.hertie-schoo.org) : Background, Unmeasured Confounding, Missing Data, Sensitivity Analysis, Analysis and Conclusion 
Ángela Duarte Pardo : Data Cleaning, G-separation, Methodology, Modelling and Analysis

# References

## Datasets
### Origital paper datasets
-   UN (2013). Population division. World Abortion policies 2013.  [Accessible here](https://www.un.org/en/development/desa/population/publications/policy/world-abortion-policies-2013.asp)

### Updated datasets
- WHO. (2022). Global Abortion Policies Worldwide. [Accessible here](https://abortion-policies.srhr.org/)

- WHO. (2022). Maternal Mortality Database.[Accessible here:](https://platform.who.int/mortality/themes/theme-details/topics/topic-details/MDB/maternal-conditions)

### Other Data sets

-	OSF. (2020). Global abortion Incidence (https://osf.io/6t4eh/)

- - WHO. (2022). Global Abortion Laws. [Accessible here](https://abortion-policies.srhr.org/)

-	Health system. (https://apps.who.int/nha/database), hospital beds (https://data.worldbank.org/indicator/SH.MED.BEDS.ZS)
-	Economic: GPD, human capital (education, as in Jensen paper), poverty (https://data.worldbank.org/indicator/SI.POV.NAHC), GINI (https://data.worldbank.org/indicator/SI.POV.GINI)

- life expectancy at birth (https://data.worldbank.org/indicator/SP.DYN.LE00.IN)

###Text